---
title: "Homework 3"
output: github_document
---

Hannah Briggs (heb2133)

```{r}
library(tidyverse)
```

## Problem 1

Load Instacart dataset.

```{r}
library(p8105.datasets)
data("instacart")
```

This is an online grocery shopping (Instacart) dataset, that can be used to explore trends in purchases. It contains `r nrow(instacart)` rows and `r ncol(instacart)` columns. Observations are the level of items in orders by customer (i.e., user). There are customer(user)/order variables -- user ID, order ID, order day, order hour, and days since prior order. In addition, there are item variables -- product ID, product name, aisle name, aisle ID, department name, and department ID. 

Looking at aisles in dataset:

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

There are `r nrow(instacart %>% count(aisle))` aisles in the dataset. The top five aisles where the most items are ordered from are: Fresh vegetables (150,609); Fresh Fruits (150,473); Packaged vegetables fruits (78,493); Yogurt (55,240); and Packaged cheese (41,699).

Plot showing number of items ordered in each aisle (limited to aisles with > 10000 items ordered):

```{r}
instacart %>% 
  count(aisle) %>%
  filter(n > 10000) %>% 
  mutate(
    aisle = factor(aisle),
    aisle = fct_reorder(aisle, n)
  ) %>% 
  ggplot(aes(x = aisle, y = n)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 270, vjust = 0.5, hjust = 1))
```

Table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”:

```{r}
instacart %>% 
  filter(aisle %in%  c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  #Get a ranking for items ordered
  filter(rank < 4) %>% 
  #Get the 3 most popular items
  arrange(aisle, rank) %>% 
  knitr::kable()
```

Table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week:

```{r}
instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = mean(order_hour_of_day)) %>% 
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hour
  ) %>% 
  rename(
    "sunday" = "0",
    "monday" = "1",
    "tuesday" = "2",
    "wednesday" = "3",
    "thursday" = "4",
    "friday" = "5",
    "saturday" = "6"
  ) %>% 
  knitr::kable()
```


## Problem 2

Load and tidy accelerometer dataset:

```{r}
accel_df = 
  read.csv("./data/accel_data.csv") %>%
  janitor::clean_names() %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_to = "minute",
    names_prefix = "activity_",
    values_to = "activity_count"
  ) %>% 
  mutate(
    day = factor(day),
    minute = as.numeric(minute),
    weekday_or_weekend = recode(day, 
                             "Monday" = "weekday",
                             "Tuesday" = "weekday",
                             "Wednesday" = "weekday",
                             "Thursday" = "weekday",
                             "Friday" = "weekday",
                             "Saturday" = "weekend",
                             "Sunday" = "weekend")
  ) %>% 
  rename("day_of_observation" = "day_id")
  
```

This dataset is generated from an accelerometer measuring the daily activity (i.e., activity measured for each minute of the day) of a 63 year-old male cardiac patient for a period of 5 weeks. Variables include week number (1-5), number of days into observation (1-35), day of the week, minute of the day, activity count, and whether observation is on a weekday or weekend. There are `r nrow(accel_df)` rows and `r ncol(accel_df)` columns in this dataset.

Trends of daily activity over five weeks:

```{r}
accel_df %>% 
  mutate(
    day = factor(day),
    day = forcats::fct_relevel(day, c("Sunday",
                                       "Monday", 
                                       "Tuesday", 
                                       "Wednesday",
                                       "Thursday",
                                       "Friday",
                                       "Saturday"))
  ) %>% 
  group_by(week, day) %>% 
  summarize(daily_activity = sum(activity_count)) %>% 
  knitr::kable()

```

Looking at this table we can observe potential trends in the patient's activity over the 5 weeks. The patient's activity on weekends seems to decline over time. For example: His daily activity counts in Week 1 for Saturday (376,254.0) and Sunday (631,105.0); compared to week 5 Saturday (1,440.0) and Sunday (138,421.0). His activity, on average, is lower in the last two weeks of observation, and was highest in the 2nd and 3rd weeks of observation. 

Plot of 24-hour activity for all 35 days of observation:

```{r}
accel_week1 =
  accel_df %>% 
    mutate(
    day = factor(day),
    day = forcats::fct_relevel(day, c("Sunday",
                                       "Monday", 
                                       "Tuesday", 
                                       "Wednesday",
                                       "Thursday",
                                       "Friday",
                                       "Saturday"))
  ) %>% 
  filter(week == 1)

accel_week2 = 
  accel_df %>% 
    mutate(
    day = factor(day),
    day = forcats::fct_relevel(day, c("Sunday",
                                       "Monday", 
                                       "Tuesday", 
                                       "Wednesday",
                                       "Thursday",
                                       "Friday",
                                       "Saturday"))
  ) %>% 
  filter(week == 2)

accel_week3 = 
  accel_df %>% 
    mutate(
    day = factor(day),
    day = forcats::fct_relevel(day, c("Sunday",
                                       "Monday", 
                                       "Tuesday", 
                                       "Wednesday",
                                       "Thursday",
                                       "Friday",
                                       "Saturday"))
  ) %>% 
  filter(week == 3)

accel_week4 = 
  accel_df %>% 
    mutate(
    day = factor(day),
    day = forcats::fct_relevel(day, c("Sunday",
                                       "Monday", 
                                       "Tuesday", 
                                       "Wednesday",
                                       "Thursday",
                                       "Friday",
                                       "Saturday"))
  ) %>% 
  filter(week == 4)

accel_week5 = 
  accel_df %>%
    mutate(
    day = factor(day),
    day = forcats::fct_relevel(day, c("Sunday",
                                       "Monday", 
                                       "Tuesday", 
                                       "Wednesday",
                                       "Thursday",
                                       "Friday",
                                       "Saturday"))
  ) %>% 
  filter(week == 5)

ggplot(data = accel_week1,
  aes(x = minute, y = activity_count, color = day)) + 
  geom_smooth(se = FALSE) +
  geom_smooth(data = accel_week2, se = FALSE) +
  geom_smooth(data = accel_week3, se = FALSE) +
  geom_smooth(data = accel_week4, se = FALSE) + 
  geom_smooth(data = accel_week5, se = FALSE) + 
  viridis::scale_color_viridis(
    name = "day",
    discrete = TRUE
  )
```

From this plot we can see that daily activity tends to peak roughly around 600-650 minutes, and minute 1250 of the day. This equates to activity peaks around roughly 10am, and 9pm. Meaning that in the early/late afternoon, the patient tends to be less active and relatively sedentary compared to the rest of the daytime. 

## Problem 3 

Load NY NOAA dataset. 

```{r}
library(p8105.datasets)
data("ny_noaa")
```

Looking at missing data:
```{r}
ny_noaa %>% count(snow = is.na(snow))
ny_noaa %>% count(prcp = is.na(prcp))
ny_noaa %>% count(snwd = is.na(snwd))
ny_noaa %>% count(tmax = is.na(tmax)) 
ny_noaa %>% count(tmin = is.na(tmin))
```

This dataset is from the NOAA National Climatic Data Center, containing data from NY state weather stations from January 1981 to December 2010. There are `r nrow(ny_noaa)` rows and `r ncol(ny_noaa)` columns in this dataset. Variables include, weather station ID, date, precipitation, snowfall and snow depth, and the max and min temperature.

The dataset is missing data for key variables to the following degree (number observations = na):
Snowfall (381,221), Snowdepth (591,786), Precipitation (145,838), Tmin (1,134,420), and Tmax (1,460,756).

Data cleaning: 
```{r}
ny_noaa_clean = 
  ny_noaa %>% 
  separate(date, c("year", "month", "day")) %>% 
  mutate(year = as.numeric(year), 
         day = as.numeric(day),
         month = recode(month, 
                        "1" = "january", 
                        "2" = "february",
                        "3" = "march",
                        "4" = "april",
                        "5" = "may",
                        "6" = "june",
                        "7" = "july",
                        "8" = "august",
                        "9" = "september",
                        "10" = "october",
                        "11" = "november",
                        "12" = "december")
         )
```


"reasonable units"-check in data descrip etc.

snowfall - count, commonly obs values

2panel plot - avg max temp; data manip step then plotting. jan and july (filter). organize data for avg max temp (group and summ), need ot grop by station, year, month, and summarize to get avg max temp. panels (facet, jan and july)

q3: pair plots together - patchwork. 
plot 1 - tmax/tmin, too much data for scatterplot, contout or bin or hex plot
plot 2 - filter snowfall, distrib plot (ridge, violin, box = by year)